---
title: "ABV vs. IBU comparison for US Craft Beers"
author: "Aaron Cattley and Bruce Lee"
date: "2/18/2019"
output: html_document
---
#Introduction: 
  # This will review the basic statistics of U.S. craft beer.  There are 2410 beers across 558 U.S. breweries that will be a part of this study. The two separate datasets Budweiser supplied were used concurrently to develop these statistics.
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, echo=TRUE}
library(tidyr)
library(ggplot2)
library(tidyverse)
```

```{r question_1 data, echo=TRUE}
## First we've brought the datasets into our Rmd and created a list of how many breweries there are by state.
fileLocation <- "https://raw.githubusercontent.com/BruceLeeSMU/CaseStudy1/master/Breweries.csv"
breweryData <- read.csv(url(fileLocation))

brewery.byState <- count(breweryData, State)
names(brewery.byState) <- c("State", "NumPerState")
head(brewery.byState)
tail(brewery.byState)

Stateplot <- ggplot(brewery.byState, aes(x=reorder(State, NumPerState), y = NumPerState, colour = State)) + geom_jitter(aes(col=State, size = NumPerState)) + geom_smooth(aes(col=State), method = "lm", se=F) + labs(title = "Number of Breweries by State", x = "State", y = "Number of Breweries")+ coord_flip()
Stateplot
```

  # We've built a list of breweries by state.  Based on this list: Colorado, California, Michigan, Oregon and Texas have the highest number of breweries.  The states with a lower number of breweries appear to be more related to being geographically based in more rural states than the size of the state.

```{r question_2 data, echo=TRUE}
## We'll use the beer data set and merge it with the brewery data set.  We will show some examples of how this combined data frame looks.
fileLocation2 <- "https://raw.githubusercontent.com/BruceLeeSMU/CaseStudy1/master/Beers.csv"
beerData <- read.csv(url(fileLocation2))

breweryAndBeerData <- merge(breweryData, beerData, by.x="Brew_ID", by.y="Brewery_id")

head(breweryAndBeerData, 6)
tail(breweryAndBeerData, 6)
```

  # The newly merged dataset is ordered by the by the Brew ID.  A sample of the beginning and ending variables are shown for example. 

```{r question_3 data, echo=TRUE}
## Data sets sometimes have NA's for various reasons; we'll extract those values for representational purposes.  
str(breweryAndBeerData)
## Rename the Name.x and Name.y columns from after merging
colnames(breweryAndBeerData)[2] <- "BreweryName"
colnames(breweryAndBeerData)[5] <- "BeerName"

summary(breweryAndBeerData)

num.NA.ABV <- length(which(is.na(breweryAndBeerData$ABV)))
num.NA.IBU <- length(which(is.na(breweryAndBeerData$IBU)))

num.NA <- data.frame(((num.NA.ABV/2410)*100),((num.NA.IBU/2410)*100))
colnames(num.NA) <- c("ABV", "IBU")
num.NA <- stack(num.NA)
num.NA <- as.vector(num.NA)

barplot(num.NA$values, ylim = c(0,50))
## From the summary of the data set, NA's are only present in the ABV and IBU columns where ABV has 62 NA values and IBU has 1005 NA values.

```
  #Due to the large dataset we are working with, the lack of available data due the NA's will not impact the statistical analysis.

```{r question_4 data, echo=TRUE}
#The next chunk of code will calculate the medians for both the ABV and IBU of all beers in the data set. Histograms will also be coded for visualization.
ABVMed <- aggregate(breweryAndBeerData$ABV ~ breweryAndBeerData$State, data = breweryAndBeerData, median)
colnames(ABVMed) <- c("State", "ABV")
ABVMed <- ABVMed[order(ABVMed$ABV, decreasing = T),]

IBUMed <- aggregate(breweryAndBeerData$IBU ~ breweryAndBeerData$State, data = breweryAndBeerData, median)
colnames(IBUMed) <- c("State", "IBU")
IBUMed <- IBUMed[order(IBUMed$IBU, decreasing = T),]

ABVplot <- ggplot(ABVMed, aes(x=reorder(State, -ABV), y = ABV, colour = State)) + geom_bar(stat = "Identity") + labs(title = "ABV Median by State", x = "State", y = "Median")+ coord_flip()
ABVplot

IBUplot <- ggplot(IBUMed, aes(x=reorder(State, -IBU), y = IBU, colour = State)) + geom_bar(stat = "Identity") + labs(title = "IBU Median by State", x = "State", y = "Median")+ coord_flip()
IBUplot
```
  # We've plotted the ABV and IBU with histograms.  They've been ordered from largest to smallest.
```{r question_5 data, echo=TRUE}
#Identify the State with the highest ABV beer
head(ABVMed, 1)
#Identify the State with the highest IBU beer
head(IBUMed, 1)
```

```{r question_6 data, echo=TRUE}
#The following code will provide a summary of statistics for the ABV.
summary(breweryAndBeerData$ABV)
```
  #The ABV has a mean of 0.05977 and median of 0.067.  This would indicate this data has a skewed distribution.  Should more advanced statistics be run, the tests should be robust due to the large sample size in accordance to the Central Limit Theorem.
```{r question_7 data, echo=TRUE}
#Finally, we will build a scatterplot of the relationship between ABV and IBU
breweryAndBeerData2 <- na.omit(breweryAndBeerData)

BandBplot <- ggplot(breweryAndBeerData2, aes(x = breweryAndBeerData2$ABV, y = breweryAndBeerData2$IBU)) + geom_point(size = 2, shape = 23) + geom_point(aes(x=breweryAndBeerData2$ABV, y=breweryAndBeerData2$IBU, color = breweryAndBeerData2$State == 'KY' | breweryAndBeerData2$State == 'DC', size = breweryAndBeerData2$State == 'KY' | breweryAndBeerData2$State == 'DC')) + 
  scale_color_manual(values = c("black", "red")) + scale_size_manual(values =c(1, 3)) + theme(legend.position = "none") +  geom_smooth(method= "lm") + labs(title = "Relationship between ABV and IBU", x = "ABV", y = "IBU") + theme(plot.title = element_text(hjust = 0.5))
BandBplot
```

  # The scatterplot shows a trend that increasing ABV in beer also increases the bitter taste.  
  
  # To strengthen our findings, a correlation test was run and there was significant evidence at the alpha level of 0.05 level of significance to suggest that the ABV and IBU are linearly correlated (p < 0.0001)!
```{r further_evidence data, echo=TRUE}
plot(breweryAndBeerData2$ABV, breweryAndBeerData2$IBU, xlab='ABV', ylab='IBU', main = 'ABV vs IBU')
cor.test(breweryAndBeerData2$ABV, breweryAndBeerData2$IBU)
```